---
title: "CA House Price Prediction"
author: Jennifer and Eli
output: html_document
---


```{r hw2_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE)
```


```{r, warning = FALSE, message = FALSE, echo = FALSE}
# library statements 
# read in data
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels)
library(readxl)
tidymodels_prefer()

housing <- read_csv("housing.csv") %>% 
  drop_na()
```


```{r, include = FALSE, warning = FALSE, message = FALSE, echo = FALSE}
# Re-coding ocean proximity to numeric values 

housing$ocean_proximity[housing$ocean_proximity=="NEAR BAY"] <- 1
housing$ocean_proximity[housing$ocean_proximity=="<1H OCEAN"] <- 2
housing$ocean_proximity[housing$ocean_proximity=="NEAR OCEAN"] <- 3
housing$ocean_proximity[housing$ocean_proximity=="ISLAND"] <- 4
housing$ocean_proximity[housing$ocean_proximity=="INLAND"] <- 5
```


```{r, creation of cv folds}
set.seed(123)
data_cv10<- vfold_cv(housing, v = 10)
```



<br><br><br>

## OLS Analysis {-}


```{r, OLS model spec}
lm_spec <-
    linear_reg() %>%
    set_engine(engine = 'lm') %>%
    set_mode('regression')
```



### OLS Recipe and Workflow Setup

```{r, OLS recipe + workflow}
lm_rec <- recipe(median_house_value ~ ., data = housing) %>%
  step_normalize(all_numeric_predictors())

lm_wf <- workflow() %>%
  add_recipe(lm_rec) %>%
  add_model(lm_spec)

housing_lm_wf_1 <- workflow() %>%
  add_recipe(lm_rec) %>%
  add_model(lm_spec)%>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_nzv(all_predictors())

housing_lm_wf_2 <- workflow() %>%
  add_formula(median_house_value ~ median_income)%>%
  add_model(lm_spec)%>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_nzv(all_predictors())

housing_lm_wf_3 <- workflow() %>%
  add_formula(median_house_value ~ median_income + longitude)%>%
  add_model(lm_spec)%>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_nzv(all_predictors())

housing_lm_wf_4 <- workflow() %>%
  add_formula(median_house_value ~ median_income + longitude + latitude)%>%
  add_model(lm_spec)%>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_nzv(all_predictors())

housing_lm_wf_5 <- workflow() %>%
  add_formula(median_house_value ~ median_income + longitude + latitude + total_rooms)%>%
  add_model(lm_spec)%>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_nzv(all_predictors())

housing_lm_wf_6 <- workflow() %>%
  add_formula(median_house_value ~ median_income + longitude + latitude + total_rooms + population)%>%
  add_model(lm_spec)%>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_nzv(all_predictors())
```


### OLS Fit and Tune Model

```{r, OLS: Fit and Tune Model}
mod1 <- fit(lm_spec,
            median_house_value ~ .,
            data = housing)
mod2 <- fit(lm_spec,
            median_house_value ~ median_income,
            data = housing)
mod3 <- fit(lm_spec,
            median_house_value ~ median_income + longitude,
            data = housing)
mod4 <- fit(lm_spec,
            median_house_value ~ median_income + longitude + latitude,
            data = housing)
mod5 <- fit(lm_spec,
            median_house_value ~ median_income + longitude + latitude + total_rooms,
            data = housing)
mod6 <- fit(lm_spec,
            median_house_value ~ median_income + longitude + latitude + total_rooms + population,
            data = housing)

mod1 %>% 
  tidy() %>% 
  slice(-1) %>% 
  mutate(lower = estimate - 1.96*std.error, upper = estimate + 1.96*std.error) %>%
  ggplot() + 
    geom_vline(xintercept=0, linetype=4) + 
    geom_point(aes(x=estimate, y=term)) + 
    geom_segment(aes(y=term, yend=term, x=lower, xend=upper), 
                 arrow = arrow(angle=90, ends='both', length = unit(0.1, 'cm'))) + 
    labs(x = 'Coefficient estimate (95% CI)', y = 'Feature') +
    theme_classic()

mod1 %>%
  tidy()
```


```{r, OLS}
mod1_output <- mod1 %>% 
    predict(new_data = housing) %>%
    bind_cols(housing) %>%
    mutate(resid = median_house_value - .pred)

mod2_output <- mod2 %>% 
    predict(new_data = housing) %>%
    bind_cols(housing) %>%
    mutate(resid = median_house_value - .pred)

mod3_output <- mod3 %>% 
    predict(new_data = housing) %>%
    bind_cols(housing) %>%
    mutate(resid = median_house_value - .pred)

mod4_output <- mod4 %>% 
    predict(new_data = housing) %>%
    bind_cols(housing) %>%
    mutate(resid = median_house_value - .pred)

mod5_output <- mod5 %>% 
    predict(new_data = housing) %>%
    bind_cols(housing) %>%
    mutate(resid = median_house_value - .pred)

mod6_output <- mod6 %>% 
    predict(new_data = housing) %>%
    bind_cols(housing) %>%
    mutate(resid = median_house_value - .pred)
```



```{r, OLS setup mae}
mod1_output %>%
    mae(truth = median_house_value, estimate = .pred)

mod2_output %>%
    mae(truth = median_house_value, estimate = .pred)

mod3_output %>%
    mae(truth = median_house_value, estimate = .pred)

mod4_output %>%
    mae(truth = median_house_value, estimate = .pred)

mod5_output %>%
    mae(truth = median_house_value, estimate = .pred)

mod6_output %>%
    mae(truth = median_house_value, estimate = .pred)
```



### OLS Visualize Model

```{r, OLS visualization with mod1 outputs}
ggplot(mod1_output, aes(y=resid, x=median_house_value)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

#not very useful?
ggplot(mod1_output, aes(y=resid, x=housing_median_age)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

#not very useful 
ggplot(mod1_output, aes(y=resid, x=ocean_proximity)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod1_output, aes(y=resid, x=longitude)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod1_output, aes(y=resid, x=latitude)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod1_output, aes(y=resid, x=total_rooms)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod1_output, aes(y=resid, x=median_income)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod1_output, aes(y=resid, x=total_bedrooms)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```


```{r, OLS visualization with mod2 outputs}
ggplot(mod2_output, aes(y=resid, x=total_rooms)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod2_output, aes(y=resid, x=median_house_value)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod2_output, aes(y=resid, x=longitude)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod2_output, aes(y=resid, x=latitude)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod2_output, aes(y=resid, x=median_income)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod2_output, aes(y=resid, x=population)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```


```{r, OLS Visualization}
ggplot(mod3_output, aes(y=resid, x=population)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod3_output, aes(y=resid, x=total_rooms)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(mod6_output, aes(y=resid, x=longitude)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```


### OLS Calculate and Collect Metrics

```{r, OLS calculate metrics}
mod1_cv <- fit_resamples(housing_lm_wf_1,
  resamples = data_cv10, 
  metrics = metric_set(rmse, rsq, mae))
  
  
mod2_cv <- fit_resamples(housing_lm_wf_2,
  resamples = data_cv10, 
  metrics = metric_set(rmse, rsq, mae))
  
  
mod3_cv <- fit_resamples(housing_lm_wf_3,
  resamples = data_cv10, 
  metrics = metric_set(rmse, rsq, mae))
  
  
mod4_cv <- fit_resamples(housing_lm_wf_4,
  resamples = data_cv10, 
  metrics = metric_set(rmse, rsq, mae))
  
  
mod5_cv <- fit_resamples(housing_lm_wf_5,
  resamples = data_cv10, 
  metrics = metric_set(rmse, rsq, mae))

  
mod6_cv <- fit_resamples(housing_lm_wf_6,
  resamples = data_cv10, 
  metrics = metric_set(rmse, rsq, mae))
```


```{r, OLS collect metrics}
mod1_cv %>% unnest(.metrics) %>%
  filter(.metric == 'rmse') %>%
  summarize(RMSE_CV = mean(.estimate))

mod2_cv %>% unnest(.metrics) %>%
  filter(.metric == 'rmse') %>%
  summarize(RMSE_CV = mean(.estimate))

mod3_cv %>% unnest(.metrics) %>%
  filter(.metric == 'rmse') %>%
  summarize(RMSE_CV = mean(.estimate))

mod4_cv %>% unnest(.metrics) %>%
  filter(.metric == 'rmse') %>%
  summarize(RMSE_CV = mean(.estimate))

mod5_cv %>% unnest(.metrics) %>%
    filter(.metric == 'rmse') %>%
    summarize(RMSE_CV = mean(.estimate))

mod6_cv %>% unnest(.metrics) %>%
    filter(.metric == 'rmse') %>%
    summarize(RMSE_CV = mean(.estimate))

mod1_cv %>% collect_metrics()
mod2_cv %>% collect_metrics()
mod3_cv %>% collect_metrics()
mod4_cv %>% collect_metrics()
mod5_cv %>% collect_metrics()
mod6_cv %>% collect_metrics()
```



<br><br><br>

## LASSO Analysis {-}

### LASSO Model spec, recipe, and workflow 

```{r, LASSO model spec}
lasso_spec <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso, we'll choose penalty later
  set_engine(engine = 'glmnet') %>% #note we are using a different engine
  set_mode('regression') 
```


```{r, lasso recipe + workflow}
lasso_rec <- recipe(median_house_value ~ ., data = housing) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value
    #step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # important standardization step for LASSO
    step_dummy(all_nominal_predictors())  # creates indicator variables for categorical variables

lasso_wf <- workflow() %>% 
  add_recipe(lasso_rec) %>%
  add_model(lasso_spec)
```


### LASSO Fit and Tune Model


```{r, LASSO: Fit and Tune Model}
# Tune LASSO #1
penalty_grid <- grid_regular(
  penalty(range = c(-5, 3)), #log10 transformed 10^-5 to 10^3
  levels = 30)

# Tune LASSO #2
tune_res <- tune_grid( # new function for tuning parameters
  lasso_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(rmse, mae),
  grid = penalty_grid)

# LASSO model
lasso_fit <- lasso_wf %>% 
  fit(data = housing) # Fit to data
```


### LASSO Calculate and Collect Metrics


```{r, calculate/collect CV metrics}
# plotting LASSO lambda
plot(lasso_fit %>% extract_fit_parsnip() %>% pluck('fit'), # way to get the original glmnet output
     xvar = "lambda")

# Visualize LASSO Metrics from Tuning
autoplot(tune_res) + theme_classic()

# Summarize LASSO CV Metrics
collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>% # or choose mae
  select(penalty, rmse = mean) 

# choose penalty value based on lowest mae or rmse
best_penalty <- select_best(tune_res, metric = 'rmse') 
best_penalty
# choose penalty value based on the largest penalty within 1 se of the lowest CV MAE
best_se_penalty <- select_by_one_std_err(tune_res, metric = 'mae', desc(penalty)) 


# Fit Final LASSO Models
final_wf <- finalize_workflow(lasso_wf, best_penalty) # incorporates penalty value to workflow
final_wf_se <- finalize_workflow(lasso_wf, best_se_penalty) # incorporates penalty value to workflow


final_fit <- fit(final_wf, data = housing)
final_fit_se <- fit(final_wf_se, data = housing)

tidy(final_fit)
tidy(final_fit_se)
```


### LASSO Visualize residuals 

```{r, LASSO visual residuals}
lasso_mod_out <- final_fit_se %>%
    predict(new_data = housing) %>%
    bind_cols(housing) %>%
    mutate(resid = median_house_value - .pred)

lasso_mod_out %>% 
  ggplot(aes(x = .pred, y = resid)) + 
  geom_point() +
  geom_smooth(se = FALSE) + 
  geom_hline(yintercept = 0) + 
  theme_classic()
```


C.Compare estimated test performance across the models. Which models(s) might you prefer?

$~$

 Using the measures calculated using the ordinary least squares. Also, using the average cross validated rmse model 1 perforomed the bed. The first model including all the variables is the best. This is also validated using the best lambda measure calculated using lasso. 
 
$~$
 
D. Use residual plots to evaluate whether some quantitative predictors might be better modeled with nonlinear relationships.

Using the residual plots, latitude and longitude might be better modeled with a nonlinear relationship. 

$~$

E. Which variables do you think are the most important predictors of your quantitative outcome? Justify your answer. Do the methods you’ve applied reach consensus on which variables are most important? What insights are expected? Surprising?

The most important predictor according to the lasso is median income. It survives the longest. median income also has the lowest p value. This is not surprising as a higher income would suggest the ability to purchase a more expensive home. Using p value population the relationship observced in population has the second lowest probability of being caused by random chance, however p value is not always the best predictor of which variables to include in the model.
Note that if some (but not all) of the indicator terms for a categorical predictor are selected in the final models, the whole predictor should be treated as selected.

Portfolio Links:
Eli: https://docs.google.com/document/d/15rt9RRehvTrdzZk_6s0136uYS9XfJcKhsydsaoY9UvE/edit?usp=sharing
Jennifer:  https://docs.google.com/document/d/1pu4LnKgN1YVL5-afdVjBd4yFJAI77m-_mTwYq3R0PB4/edit?usp=sharing



